Left-linear rules in adhesive categories

### 2024-02-05 icalp/fusione.tex
app

\section{$\mathcal{M}$-adhesive categories and DPO rewriting systems}\label{sec:ade}
\todo{RIDURRE MOLTO}

@@

\begin{definition}[Pushout complement]
Let $f\colon X\to Y$ and $g\colon Y\to Z$ ba two composable arrows in a category $\X$. A \emph{pushout complement} for the pair $(f,g)$ is a pair $(h,k)$ with $h\colon X\to W$ and $k\colon W\to Z$ such that the square below commutes and it is a pushout.
\[\xymatrix{X \ar[r]^{f} \ar[d]_{h}& Y \ar[d]^{g} \\ W \ar[r]_{k}& Z}\]
\end{definition}
\iffalse 
\begin{example}
	In a generic category $\X$, pushout complements may not exist: in $\Set$ the arrows $?_{2}\colon \emptyset \to 2$ and $!_2\colon 2\to 1$ do not have a pushout complement.

@@

\begin{theoremEnd}[category=sec2]{lemma}\label{lem:radj}
Let $f\colon X\to Y$ be an arrow in an $\mathcal{M}$-adhesive category $\X$ and suppose that the left square below is a pushout while the  right one is a pullback, with $ m\colon M\to X$ and $n\colon N\to Y$ in $\mathcal{M}$.
\[\xymatrix{M \ar@{>->}[d]_{m}\ar[r]^{h}& Q \ar@{>->}[d]^{k} & P \ar[r]^{p_1} \ar@{>->}[d]_{p_2}& N \ar@{>->}[d]^{n}\\X\ar[r]_{f}&Y &X \ar[r]_f&Y\\}\]
Then $n\leq k$ if and only if $p_2\leq m$.
\end{theoremEnd}
\begin{proof} 
\begin{proofEnd}  
	We start pulling back $n$ along $k$ and then pulling back the resulting arrow along $h$, getting the following two pullbacks square:
\[\xymatrix{B \ar@{>->}[d]_{b}\ar@{>->}[r]^{s}& N \ar@{>->}[d]^{n} & A \ar[r]^{r} \ar@{>->}[d]_{a}& B \ar@{>->}[d]^{b}\\Q\ar@{>->}[r]_{k}&Y &M \ar[r]_h&Q\\}\]
Notice that 
\[
f\circ m\circ a=k\circ h\circ a=k\circ b\circ r=n\circ s\circ r\]

@@

	The thesis now follows immediately from \Cref{lem:varie}. \qedhere 
\end{proofEnd}
\end{proof}
\begin{theoremEnd}[category=sec2]{corollary}[Uniqueness of pushouts complements]\label{lem:pocomp} 
Let $\X$ be a $\mathcal{M}$-adhesive category. Given $m\colon X\mto Y$ in $\mathcal{M}$ and $f\colon Y\to Z$, let $(h_1, k_1)$ and $(h_2, k_2)$ be pushout complements of $m$  and $f$ and $W_1=\cod{h_1}$, $W_2=\cod{h_2}$. Then there exists a unique isomorphism $\phi\colon W_1\to W_2$ making the following diagram commutative.
\[\xymatrix{&X \ar@{>->}[r]^{m} \ar[d]_{h_1} \ar@/_.3cm/[dl]_{h_2}& Y \ar[d]^{f} \\ W_2 \ar@{>->}@/_.4cm/[rr]_{k_2} &W_1 \ar@{.>}[l]_{\phi} \ar@{>->}[r]^{k_1} & Z }\]
\end{theoremEnd}
\begin{proof}\begin{proofEnd}
	Consider the two pushout squares
	\[\xymatrix{M \ar[r]^{h_1}\ar@{>->}[d]_{m}& W_1 \ar@{>->}[d]^{k_1} & M \ar@{>->}[d]_{m} \ar[r]^{h_1}& W_2 \ar@{>->}[d]^{k_2}\\Y\ar[r]_{f}&Z &Y \ar[r]_{f}&Z}\]
	Since $\mathcal{M}$ is in $\mathcal{M}$, \cref{prop:pbpoad} guarantees that they are both pullbacks.
	Since $m\leq m$, by \Cref{lem:radj} we get that now entails that $k_1\leq k_2$ and $k_2\leq k_1$. Thus there exists an isomorphism $\phi\colon W_1\to W_2$ such that $k_1=k_2\circ \phi$. To see that $h_2=\phi\circ h_1$, we can compute:
	\[
	k_2\circ \phi \circ h_1  = k_1\circ h_1= n\circ m= k_2\circ h_2\]

@@

\begin{definition}[\cite{habel2012mathcal,heindel2009category}]
	Let $\X$ be a $\mathcal{M}$-adhesive category, a  \emph{left $\mathcal{M}$-linear} rule $\rho$ is a pair $(l,r)$ of arrows with the same domain, such that $l$ belongs to $\mathcal{M}$.  The rule $\rho$ is said to be \emph{$\mathcal{M}$-linear} if $r\in \mathcal{M}$ too. A rule $\rho$ is said to be \emph{consuming} if $l$ is not an isomorphism. We will represent a rule $\rho$ as a span 
	\[\xymatrix{L & K\ar[l]_{l} \ar[r]^{r} & R}\]
	$L$ is the \emph{left-hand side}, $R$ is the \emph{right-hand side} and $K$ the \emph{glueing object}.

@@

	Given  two objects $G$ and $H$ and a rule $\rho=(l,r)$ in $\R$, a \emph{direct derivation $\mathscr{D}$ from $G$ to $H$ applying the rule $\rho$}, is a diagram as the one below, in which both squares are pushouts. 
	\[\xymatrix{L \ar[d]_{n}& K \ar[d]^{k}\ar[l]_{l} \ar[r]^{r} & R \ar[d]^{h}\\G & \ar[l]^{f} D \ar[r]_{g}& H}\]
	The arrow $n$ is called the \emph{match} of the derivation, while $h$ is its \emph{back-match}.
	We will denote a direct derivation $\dder{D}$ between $G$ and $H$ as $\dder{D}\colon G\Mapsto H$. 
\end{definition}

\begin{theoremEnd}[category=sec2]{proposition}\label{prop:unique} Let $\X$  be a $\mathcal{M}$-adhesive category. Suppose that the two direct derivations $\mathscr{D}$ and $\mathscr{D'}$ below, with the same match and applying the same left-linear rule $\rho$ are given.
	\[\xymatrix{L \ar[d]_{m}& K \ar[d]^{k}\ar@{>->}[l]_{l} \ar[r]^{r} & R \ar[d]^{h} & L \ar[d]_{m}& K \ar[d]^{k'}\ar@{>->}[l]_{l} \ar[r]^{r} & R \ar[d]^{h'}\\G & \ar@{>->}[l]^{f} D \ar[r]_{g}& H & G & \ar@{>->}[l]^{f'} D' \ar[r]_{g'}& H'}\]
	Then there exist isomorphisms $t\colon D\to D'$ and $s\colon H\to H'$ as in the following diagram.
	\[\xymatrix@C=40pt{G&D' \ar[r]^{g'} \ar@{>->}[l]_{f'}&H'\\ L \ar[d]^{m} \ar[u]_{m} & K \ar[u]_{k'} \ar[d]^{k}\ar[r]^{r} \ar@{>->}[l]_{l} &R\ar[u]^{h'} \ar[d]_{h}\\G\ar@/^.4cm/[uu]^{\id{G}}&D\ar@{>->}[l]^{f}\ar@/^.4cm/@{.>}[uu]^(.4){\phi}|\hole \ar[r]_{g}&H\ar@/_.4cm/@{.>}[uu]_{\psi}}\]
\end{theoremEnd}
\begin{proof}\begin{proofEnd}
	By construction, the pairs $(k, f)$ and $(k', f')$ are pushout complements of $l$ and $n$. Thus, the existence of the isomorphism $t\colon D\to D'$ follows from \Cref{lem:pocomp}. Now, computing we have
\[
g'\circ t \circ k= g' \circ k'=h'\circ r\]
	Hence, we have the wanted $s\colon H\to H'$. To see that $s$ is an isomorphism, consider the diagram 
	\[\xymatrix{K  \ar@/^.4cm/[rr]^{k'}\ar[d]_{r} \ar[r]_{k}& \ar[r]_{t} D \ar[d]^{g}& D' \ar[d]^{g'}\\ R \ar@/_.4cm/[rr]_{h'} \ar[r]^{h}& H \ar[r]^{s}& H'}\]
	By hypothesis the whole rectangle and its left half are pushouts, therefore, by \Cref{lem:po1} its right square is a pushout too. The claim now follows from the fact that the pushout of an isomorphism is an isomorphism.
	\end{proofEnd}
\end{proof}

@@

\begin{notation}Let $\der{D}=\{\dder{D}_i\}_{i=0}^n$ be a derivation. We will depict the $i^\text{th}$ element $\dder{D}_i$ of $\der{D}$ as in the following diagram.  
	\[\xymatrix{L_i \ar[d]_{m_i}& K_i \ar[d]^{k_i}\ar[l]_{l_i} \ar[r]^{r_i} & R_i \ar[d]^{h_i} \\G_{i} & \ar[l]^{f_{i}} D_{i} \ar[r]_{g_{i}}& G_{i+1} }\]
	Notice that, in particular, if $\der{D}\colon G\to H$, then $G_0=G$ and $G_{n+1}=H$. When $\der{D}$ has length $1$ we will suppress the indexes. In such case, we will also identify $\der{D}$ with its only element. 
\end{notation}

@@

\begin{definition}Let $(\X, \R)$ be a left-linear DPO rewriting system,  an \emph{abstraction equivalence} between two derivations $\der{D}$ and $\der{D'}$ with the same length and such that $r(\der{D})=r(\der{D}')$, is a family of isomorphisms $\{\phi_X\}_{X\in \Deltamin(\der{D})}$ such that, for every $i\in [0, \lgh(\der{D})]$ the following diagram commutes
	\[\xymatrix@C=40pt{G'_i&D'_i \ar[r]^{g'_i} \ar[l]_{f'_i}&G'_{i+1}\\  L_i \ar[u]^{n'_i} \ar[d]_{n_i}& K_i \ar[u]^{k'_i} \ar[d]_{k_i} \ar[r]^{r_i} \ar[l]_{l_i} &R\ar[u]^{h'_i} \ar[d]_{h_i}\\G_i \ar@/_.45cm/[uu]_(.35){\phi_{G_i}}|\hole&D_i\ar[l]^{f_i}\ar@/_.45cm/[uu]_(.35){\phi_{D_i}}|\hole \ar[r]_{g_i}&G_{i+1}\ar@/_.45cm/[uu]_{\phi_{H_i}}}\]

We will say that  $\der{D}$ $\der{D}'$  are \emph{abstraction equivalent}, if there exists an of abstraction equivalence between them.
\end{definition}

@@

 Let $(i_0, i_1)$ and $(i'_0, i'_1)$ be independence pairs for the direct derivations $\dder{D}_0$ and $\dder{D}_1$. Notice that, by definition, we have
 \[f_1\circ i_0=h_0=f_1\circ i'_0\]
	On the other hand, 
	$f_1\colon D_\to G_1$ is  the pushout of $l_1\colon K_1\to L_1$ and so it is in $\mathcal{M}$, implying $i_0=i'_0$. If, moreover, we suppose that the rule $\rho$ applied in $\dder{D}$ is linear, then $g_0\colon D_0\to G_1$ is in $\mathcal{M}$ too, hence, from the equation
	\[
	g_0\circ i_1=m_1= g_0\circ i'_1\]
	we can deduce that $i'_0=i'_1$, too. Summing up, we have just proved the following.

	\begin{proposition}\label{rem:weak}
	Let $(\X, \R)$ be a linear DPO rewriting system, then two direct derivations $\dder{D}$ and $\dder{D}'$ are weakly sequential independet if and only if they are sequential independent. 
\end{proposition}

@@

\begin{theoremEnd}[category=sec3]{proposition}\label{prop:tec}
	Let $(i_0, i_1)$ be an independence pair between two direct derivations $\dder{D}_0$ and $\dder{D}_1$ in a left-linear DPO rewriting system $(\X,\R)$. Consider the pullback of $f_1\colon D_1\to G_1$ along $g_0\colon D_0\to G_1$ (first square below),  then there exist the arrows $u_1\colon K_1\to P$ and $u_0\colon K_0\to P$ fitting in the central and right square, Moreover, the right one is always a pullback.
	\[\xymatrix{P \ar@{>->}[r]^{p_0} \ar[d]_{p_1} & D_0 \ar[d]^{g_0}& K_0 \ar[r]^{r_0}  \ar@{.>}[d]_{u_0}& R_0 \ar[d]^{i_0} & K_1 \ar@{>->}[r]^{l_1}  \ar@{.>}[d]_{u_1}& L_1 \ar[d]^{i_1}\\  D_1 \ar@{>->}[r]_{f_1} &G_1 &P \ar[r]_{p_1} & D_1 &P \ar[r]_{p_0}  & D_0}\]
\end{theoremEnd}
\begin{proof}
	\begin{proofEnd}
We start noticing that
\[
f_1\circ i_0\circ r_0=h_0\circ r_0=g_0\circ k_0 \qquad g_0\circ i_1\circ l_1=m_1\circ l_1= f_1 \circ k_1 \] 
Thus there exists the dotted $u_0\colon K_0\to P$, $u_1\colon K_1\to P$  in the diagram
\[\xymatrix{K_0 \ar[d]_{r_0} \ar@{.>}[r]_{u_0}\ar@/^.4cm/[rr]^{k_0} &P \ar@{>->}[r]_{p_0} \ar[d]_{p_1}& D_0 \ar[d]^{g_0}&K_1 \ar@{>->}[d]_{l_1} \ar@{.>}[r]_{u_1}\ar@/^.4cm/[rr]^{k_1} &P \ar[r]_{p_1} \ar@{>->}[d]_{p_0}& D_1 \ar@{>->}[d]^{f_1}\\R_0 \ar@/_.4cm/[rr]_{h_0}\ar[r]^{i_0}& D_1 \ar@{>->}[r]^{f_1}& G_1&L_1 \ar@/_.4cm/[rr]_{m_1}\ar[r]^{i_1}& D_0 \ar[r]^{g_0}& G_1}\]
By \Cref{prop:pbpoad,lem:pb1} we get that the left half of the second rectangle is a pullback. 
	\end{proofEnd} \qedhere 
\end{proof}

@@

\begin{definition}[Switch]\label{def:switc}
  \label{de:switch}
 Let $(\X, \R)$ be a left-linear DPO rewriting system with $\X$ an $\mathcal{M}$-adhesive category and suppose that $\der{D}=\{\dder{D}_i\}_{i=0}^1$  is a derivation of length $2$ made by the following two weakly sequentially independent derivations $\dder{D}_0$, $\der{D}_1$, with independence pair $(i_0, i_1)$:
 	\[\xymatrix@C=15pt{L_0 \ar[d]_{m_0}&& K_0 \ar[d]_{k_0}\ar@{>->}[ll]_{l_0} \ar[r]^{r_0} & R_0 \ar@/^.35cm/[drrr]|(.3)\hole_(.4){i_0} \ar[dr]|(.3)\hole_{h_0} && L_1 \ar@/_.35cm/[dlll]^(.4){i_1} \ar[dl]|(.3)\hole^{m_1}& K_1 \ar[d]^{k_1}\ar@{>->}[l]_{l_1} \ar[rr]^{r_1} && R_1 \ar[d]^{h_1}\\G_0 && \ar@{>->}[ll]^{f_0} D_0 \ar[rr]_{g_0}&& G_1  && \ar@{>->}[ll]^{f_1} D_1 \ar[rr]_{g_1}&& G_2}\]

@@

	 	On the other hand, the rule applied by $\dder{D}_1$ cannot be applied to $c$: indeed the unique morphism $a\to c$ is $(a,c)$, thus we get a diagram
	 	\[\xymatrix{a  \ar[d]_{(a,c)}& a \ar[d]_{(a,c)}\ar@{>->}[l]_{\id{a}} \ar[r]^{(a,b)} & b \\c & c \ar[l]^{\id{c}} }\]
	 	But $b$ and $c$ do not have a supremum in the poset underlying $\X$, thus the previous diagram cannot be completed to a direct derivation.
\end{example}

@@

\todo{chiusura per certe costruzioni}

The following result shows that the mixed and pushout decomposition properties guarantee that every independence pair is good.

@@

We have to show that $(i_0, i_1)$ is a good independence pair. Now, by hypothesis $\X$ has all pushouts, thus the only thing to show is that the squares below are pushouts (the third one is the usual pullback of $f_1\colon D_1\to G_0$ along $g_0\colon D_0\to G_1$).
			\[\xymatrix{K_0 \ar[r]^{r_0}  \ar[d]_{u_0}& R_0 \ar[d]^{i_0} & K_1 \ar@{>->}[r]^{l_1}  \ar[d]_{u_1}& L_1 \ar[d]^{i_1}&P \ar@{>->}[r]^{p_0} \ar[d]_{p_1}& D_0\ar@{.>}[d]^{g_0} \\ P \ar[r]_{p_1} & D_1 &P \ar[r]_{p_0}  & D_0&D_1 \ar@{>->}[r]_{f_1} & G_1}\]

To see this, consider the following two diagrams.
\[\xymatrix{K_0 \ar@/^.4cm/[rr]^{k_0} \ar[d]_{r_0} \ar[r]_{u_0}& \ar@{>->}[r]_{p_0} P \ar[d]_{i_0}& D_0 \ar[d]^{g_0}\\ R_0 \ar@/_.4cm/[rr]_{h_0}  \ar[r]_{u_0}\ar[r]^{i_0}& D_1 \ar@{>->}[r]^{f_1}& G_1}\]

The thesis now follows from the mixed and the $\mathcal{M}$-pushout decomposition property.
\end{proof}

@@

\begin{definition}
	\todo{classe B }
\end{definition}

>>>>

\section{$\mathcal{M}$-adhesive categories and DPO rewriting systems}\label{sec:ade}

@@

\noindent 
\parbox{11.7cm}{
\begin{definition}[Pushout complement]
Let $f\colon X\to Y$ and $g\colon Y\to Z$ be two composable arrows in a category $\X$. A \emph{pushout complement} for the pair $(f,g)$ is a pair $(h,k)$ with $h\colon X\to W$ and $k\colon W\to Z$ such that the square below commutes and it is a pushout.
\end{definition}} \quad 
\parbox{4cm}{\vspace{-.15cm}
	$\xymatrix{X \ar[r]^{f} \ar[d]_{h}& Y \ar[d]^{g} \\ W \ar[r]_{k}& Z}$}

\iffalse 
\begin{example}
	In a generic category $\X$, pushout complements may not exist: in $\Set$ the arrows $?_{2}\colon \emptyset \to 2$ and $!_2\colon 2\to 1$ do not have a pushout complement.

@@

\noindent
\parbox{9cm}{
\begin{theoremEnd}[category=sec2]{lemma}\label{lem:radj}
Let $f\colon X \to Y$ be an arrow in an $\mathcal{M}$-adhesive category $\X$ and suppose that the left square aside is a pushout while the  right one is a pullback, with $ m\colon M \rightarrowtail X$ and $n\colon N \rightarrowtail Y$ in $\mathcal{M}$.
Then $n\leq k$ if and only if $p_2\leq m$.
\end{theoremEnd}}
\parbox{3cm}{\vspace{-0ex}
	$\xymatrix{M \ar@{>->}[d]_{m}\ar[r]^{h}& Q \ar@{>->}[d]^{k} & P \ar[r]^{p_1} \ar@{>->}[d]_{p_2}& N \ar@{>->}[d]^{n}\\X\ar[r]_{f}&Y &X \ar[r]_f&Y}$}

\begin{proof} 
\begin{proofEnd}  
	We start pulling back $n$ along $k$ and then pulling back the resulting arrow along $h$, getting the following two pullbacks square:
\[\xymatrix{B \ar@{>->}[d]_{b}\ar@{>->}[r]^{s}& N \ar@{>->}[d]^{n} & A \ar[r]^{r} \ar@{>->}[d]_{a}& B \ar@{>->}[d]^{b}\\Q\ar@{>->}[r]_{k}&Y &M \ar[r]_h&Q\\}\]
Notice that 
\[
f\circ m\circ a=k\circ h\circ a=k\circ b\circ r=n\circ s\circ r\]

@@

	The thesis now follows immediately from \Cref{lem:varie}. \qedhere 
\end{proofEnd}
\end{proof}

\noindent
\parbox{10cm}{ 
\begin{theoremEnd}[category=sec2]{corollary}[Uniqueness of pushouts complements]\label{lem:pocomp} 
Let $\X$ be a $\mathcal{M}$-adhesive category. Given $m\colon X\mto Y$ in $\mathcal{M}$ and $f\colon Y\to Z$, let $(h_1, k_1)$ and $(h_2, k_2)$ be the pushout complements of $m$  and $f$ depicted aside. Then there exists a unique isomorphism $\phi\colon W_1\to W_2$ making the diagram on the right commutative.
\end{theoremEnd}}
\parbox{3cm}{$\xymatrix{&X \ar@{>->}[r]^{m} \ar[d]_{h_1} \ar@/_.3cm/[dl]_{h_2}& Y \ar[d]^{f} \\ W_2 \ar@{>->}@/_.4cm/[rr]_{k_2} &W_1 \ar@{.>}[l]_{\phi} \ar@{>->}[r]^{k_1} & Z }$}

\begin{proof}\begin{proofEnd}
	Consider the two pushout squares
	\[\xymatrix{M \ar[r]^{h_1}\ar@{>->}[d]_{m}& W_1 \ar@{>->}[d]^{k_1} & M \ar@{>->}[d]_{m} \ar[r]^{h_1}& W_2 \ar@{>->}[d]^{k_2}\\Y\ar[r]_{f}&Z &Y \ar[r]_{f}&Z}\]
	Since $\mathcal{M}$ is in $\mathcal{M}$, \cref{prop:pbpoad} guarantees that they are both pullbacks.
	Since $m\leq m$, by \Cref{lem:radj} we get that now entails that $k_1\leq k_2$ and $k_2\leq k_1$. Thus there exists an isomorphism $\phi\colon W_1\to W_2$ such that $k_1=k_2\circ \phi$. To see that $h_2=\phi\circ h_1$, we can compute:
	\[
	k_2\circ \phi \circ h_1  = k_1\circ h_1= n\circ m= k_2\circ h_2\]

@@

\begin{definition}[\cite{habel2012mathcal,heindel2009category}]
	Let $\X$ be a $\mathcal{M}$-adhesive category, a  \emph{left $\mathcal{M}$-linear} rule $\rho$ is a pair $(l,r)$ of arrows with the same domain, such that $l$ belongs to $\mathcal{M}$.  The rule $\rho$ is said to be \emph{$\mathcal{M}$-linear} if $r\in \mathcal{M}$ too. A rule $\rho$ is said to be \emph{consuming} if $l$ is not an isomorphism. We will represent a rule $\rho$ as a span 
	\[\xymatrix{L & K\ar@{>->}[l]_{l} \ar[r]^{r} & R}\]
	$L$ is the \emph{left-hand side}, $R$ is the \emph{right-hand side} and $K$ the \emph{glueing object}.

@@

	Given  two objects $G$ and $H$ and a rule $\rho=(l,r)$ in $\R$, a \emph{direct derivation $\mathscr{D}$ from $G$ to $H$ applying the rule $\rho$}, is a diagram as the one below, in which both squares are pushouts. 
	\[\xymatrix{L \ar[d]_{m}& K \ar[d]^{k}\ar@{>->}[l]_{l} \ar[r]^{r} & R \ar[d]^{h}\\G & \ar@{>->}[l]^{f} D \ar[r]_{g}& H}\]
	The arrow $n$ is called the \emph{match} of the derivation, while $h$ is its \emph{back-match}.
	We will denote a direct derivation $\dder{D}$ between $G$ and $H$ as $\dder{D}\colon G\Mapsto H$. 
\end{definition}

@@

\begin{notation}Let $\der{D}=\{\dder{D}_i\}_{i=0}^n$ be a derivation. We will depict the $i^\text{th}$ element $\dder{D}_i$ of $\der{D}$ as in the following diagram.  
	\[\xymatrix{L_i \ar[d]_{m_i}& K_i \ar[d]^{k_i}\ar@{>->}[l]_{l_i} \ar[r]^{r_i} & R_i \ar[d]^{h_i} \\G_{i} & \ar@{>->}[l]^{f_{i}} D_{i} \ar[r]_{g_{i}}& G_{i+1} }\]
	Notice that, in particular, if $\der{D}\colon G\to H$, then $G_0=G$ and $G_{n+1}=H$. When $\der{D}$ has length $1$ we will suppress the indexes. In such case, we will also identify $\der{D}$ with its only element. 
\end{notation}

@@

\noindent 
\parbox{8cm}{
\begin{definition}Let $(\X, \R)$ be a left-linear DPO rewriting system,  an \emph{abstraction equivalence} between two derivations $\der{D}$ and $\der{D'}$ with the same length and such that $r(\der{D})=r(\der{D}')$, is a family of isomorphisms $\{\phi_X\}_{X\in \Deltamin(\der{D})}$ such that, for every $i\in [0, \lgh(\der{D})]$ the diagram on the right commutes. We will say that  $\der{D}$are $\der{D}'$  are \emph{abstraction equivalent}, if there exists an abstraction equivalence between them.
\end{definition}}
\parbox{4cm}{\vspace{-1.35em}$\xymatrix@C=40pt{G'_i&D'_i \ar[r]^{g'_i} \ar@{>->}[l]_{f'_i}&G'_{i+1}\\  L_i \ar[u]^{m'_i} \ar[d]_{m_i}& K_i \ar[u]^{k'_i} \ar[d]_{k_i} \ar[r]^{r_i} \ar@{>->}[l]_{l_i} &R\ar[u]^{h'_i} \ar[d]_{h_i}\\G_i \ar@/_.45cm/[uu]_(.35){\phi_{G_i}}|\hole&D_i\ar@{>->}[l]^{f_i}\ar@/_.45cm/[uu]_(.35){\phi_{D_i}}|\hole \ar[r]_{g_i}&G_{i+1}\ar@/_.45cm/[uu]_{\phi_{H_i}}}$}

\begin{theoremEnd}[category=sec2]{proposition}\label{prop:unique} Let $\X$  be a $\mathcal{M}$-adhesive category. Suppose that the two direct derivations $\dder{D}$ and $\dder{D}'$ below, with the same match and applying the same left-linear rule $\rho$ are given.
	\[\xymatrix{L \ar[d]_{m}& K \ar[d]^{k}\ar@{>->}[l]_{l} \ar[r]^{r} & R \ar[d]^{h} & L \ar[d]_{m}& K \ar[d]^{k'}\ar@{>->}[l]_{l} \ar[r]^{r} & R \ar[d]^{h'}\\G & \ar@{>->}[l]^{f} D \ar[r]_{g}& H & G & \ar@{>->}[l]^{f'} D' \ar[r]_{g'}& H'}\]
	Then there an abstraction equivalence between $\dder{D}$ and $\dder{D}'$, whose first component is $\id{G}$.
\end{theoremEnd}
\begin{proof}\begin{proofEnd}
		Both pairs $(k, f)$ and $(k', f')$ are pushout complements of $l$ and $n$, thus, \Cref{lem:pocomp} yields an isomorphism $\phi_D\colon D\to D'$. Computing we have
		\[
		g'\circ \phi_D \circ k= g' \circ k'=h'\circ r\]

		\noindent
		\parbox{3cm}{
			$\xymatrix{K  \ar@/^.4cm/[rr]^{k'}\ar[d]_{r} \ar[r]_{k}& \ar[r]_{\phi_D} D \ar[d]^{g}& D' \ar[d]^{g'}\\ R \ar@/_.4cm/[rr]_{h'} \ar[r]^{h}& H \ar[r]^{\phi_H}& H'}$} \hfill 
		\parbox{10cm}{
		\hspace{15pt}Hence, we have another $\phi_H\colon H\to H'$. To see that $\phi_H$ is an isomorphism, consider the diagram aside. By hypothesis the whole rectangle and its left half are pushouts, therefore, by \Cref{lem:po1} its right square is a pushout too. The claim now follows from the fact that the pushout of an isomorphism is an isomorphism. \qedhere }
	\end{proofEnd}
\end{proof}

@@

 Let $(i_0, i_1)$ and $(i'_0, i'_1)$ be independence pairs for the direct derivations $\dder{D}_0$ and $\dder{D}_1$. Notice that, by definition, we have
 \[f_1\circ i_0=h_0=f_1\circ i'_0 \qquad 	g_0\circ i_1=m_1= g_0\circ i'_1\]

The arrow $f_1\colon D\to G_1$ is  the pushout of $l_1\colon K_1\mto L_1$, thus it is in $\mathcal{M}$ implying $i_0=i'_0$. If, moreover, we suppose that the rule $\rho$ applied in $\dder{D}$ is linear, then $g_0\colon D_0\to G_1$ is in $\mathcal{M}$ too entailing $i'_1=i'_1$, too. Summing up, we have just proved the following.

	\begin{proposition}\label{rem:weak}
	Let $(\X, \R)$ be a linear DPO rewriting system, then two direct derivations $\dder{D}$ and $\dder{D}'$ are weakly sequential independent if and only if they are sequential independent. 
\end{proposition}

@@

\begin{theoremEnd}[category=sec3]{proposition}\label{prop:tec}
	Let $(i_0, i_1)$ be an independence pair between two direct derivations $\dder{D}_0$ and $\dder{D}_1$ in a left-linear DPO rewriting system $(\X,\R)$. Consider the pullback of $f_1\colon D_1\mto G_1$ along $g_0\colon D_0\to G_1$ (first square below),  then there exist the arrows $u_1\colon K_1\to P$ and $u_0\colon K_0\to P$ fitting in the central and right square, Moreover, the right one is always a pullback.
	\[\xymatrix{P \ar@{>->}[r]^{p_0} \ar[d]_{p_1} & D_0 \ar[d]^{g_0}& K_0 \ar[r]^{r_0}  \ar@{.>}[d]_{u_0}& R_0 \ar[d]^{i_0} & K_1 \ar@{>->}[r]^{l_1}  \ar@{.>}[d]_{u_1}& L_1 \ar[d]^{i_1}\\  D_1 \ar@{>->}[r]_{f_1} &G_1 &P \ar[r]_{p_1} & D_1 &P \ar[r]_{p_0}  & D_0}\]
\end{theoremEnd}
\begin{proof}
	\begin{proofEnd}
We start noticing that
\[
f_1\circ i_0\circ r_0=h_0\circ r_0=g_0\circ k_0 \qquad g_0\circ i_1\circ l_1=m_1\circ l_1= f_1 \circ k_1 \] 
Thus there exists the dotted $u_0\colon K_0\to P$, $u_1\colon K_1\to P$  in the diagram
\[\xymatrix{K_0 \ar[d]_{r_0} \ar@{.>}[r]_{u_0}\ar@/^.4cm/[rr]^{k_0} &P \ar@{>->}[r]_{p_0} \ar[d]_{p_1}& D_0 \ar[d]^{g_0}&K_1 \ar@{>->}[d]_{l_1} \ar@{.>}[r]_{u_1}\ar@/^.4cm/[rr]^{k_1} &P \ar[r]_{p_1} \ar@{>->}[d]_{p_0}& D_1 \ar@{>->}[d]^{f_1}\\R_0 \ar@/_.4cm/[rr]_{h_0}\ar[r]^{i_0}& D_1 \ar@{>->}[r]^{f_1}& G_1&L_1 \ar@/_.4cm/[rr]_{m_1}\ar[r]^{i_1}& D_0 \ar[r]^{g_0}& G_1}\]
By \Cref{prop:pbpoad,lem:pb1} we get that the left half of the second rectangle is a pullback. 
	\end{proofEnd} \qedhere 
\end{proof}

@@

\begin{definition}[Switch]\label{def:switc}
  \label{de:switch}
 Let $(\X, \R)$ be a left-linear DPO rewriting system with $\X$ an $\mathcal{M}$-adhesive category and suppose that $\der{D}=\{\dder{D}_i\}_{i=0}^1$  is a derivation of length $2$ made by two weakly sequentially independent derivations $\dder{D}_0$, $\dder{D}_1$, with independence pair $(i_0, i_1)$:
 	\[\xymatrix@C=15pt{L_0 \ar[d]_{m_0}&& K_0 \ar[d]_{k_0}\ar@{>->}[ll]_{l_0} \ar[r]^{r_0} & R_0 \ar@/^.35cm/[drrr]|(.3)\hole_(.4){i_0} \ar[dr]|(.3)\hole_{h_0} && L_1 \ar@/_.35cm/[dlll]^(.4){i_1} \ar[dl]|(.3)\hole^{m_1}& K_1 \ar[d]^{k_1}\ar@{>->}[l]_{l_1} \ar[rr]^{r_1} && R_1 \ar[d]^{h_1}\\G_0 && \ar@{>->}[ll]^{f_0} D_0 \ar[rr]_{g_0}&& G_1  && \ar@{>->}[ll]^{f_1} D_1 \ar[rr]_{g_1}&& G_2}\]

@@

	 	\noindent 
	 	\parbox{10cm}{
	 	\hspace{15pt}On the other hand, the rule applied by $\dder{D}_1$ cannot be applied to $c$. indeed the unique morphism $a\to c$ is $(a,c)$, thus we get a diagram aside. But $b$ and $c$ do not have a supremum in the poset underlying $\X$, thus we do not get a direct derivation from $c$.}
	 	\parbox{3cm}{ \vspace{-.3cm}$\xymatrix{a  \ar[d]_{(a,c)}& a \ar[d]_{(a,c)}\ar@{>->}[l]_{\id{a}} \ar[r]^{(a,b)} & b \\c & c \ar[l]^{\id{c}} }$}

\end{example}

@@

\begin{lemma}\label{lem:closed}
	\todo{chiusura per certe costruzioni}
\end{lemma}
\begin{proof}
	contenuto...
\end{proof}
The following result shows that the mixed and pushout decomposition properties guarantee that every independence pair is good.

@@

We have to show that $(i_0, i_1)$ is a good independence pair. Now, by hypothesis $\X$ has all pushouts, thus the only thing to show is that the squares below are pushouts (the third one is the usual pullback of $f_1\colon D_1\mto G_0$ along $g_0\colon D_0\to G_1$).
			\[\xymatrix{K_0 \ar[r]^{r_0}  \ar[d]_{u_0}& R_0 \ar[d]^{i_0} & K_1 \ar@{>->}[r]^{l_1}  \ar[d]_{u_1}& L_1 \ar[d]^{i_1}&P \ar@{>->}[r]^{p_0} \ar[d]_{p_1}& D_0\ar[d]^{g_0} \\ P \ar[r]_{p_1} & D_1 &P \ar[r]_{p_0}  & D_0&D_1 \ar@{>->}[r]_{f_1} & G_1}\]

To see this, consider the following two diagrams.
\[\xymatrix{K_0 \ar@/^.4cm/[rr]^{k_0} \ar[d]_{r_0} \ar[r]_{u_0}& \ar@{>->}[r]_{p_0} P \ar[d]_{p_1}& D_0 \ar[d]^{g_0}&K_1 \ar@/^.4cm/[rr]^{k_1} \ar@{>->}[d]_{l_1} \ar[r]_{u_1}& \ar[r]_{p_1} P \ar@{>->}[d]_{p_0}& D_1 \ar@{>->}[d]^{f_1}\\ R_0 \ar@/_.4cm/[rr]_{h_0}  \ar[r]^{i_0}& D_1 \ar@{>->}[r]^{f_1}& G_1 &L_1 \ar@/_.4cm/[rr]_{m_1}  \ar[r]^{i_1}& D_0 \ar[r]^{g_0}& G_1}\]

The thesis now follows from the $\mathcal{M}$-mixed and the $\mathcal{M}$-pushout decomposition property.
\end{proof}

@@

\begin{definition}
We will say that an $\mathcal{M}$-adhesive category $\X$ is of \emph{type $\mathbb{B}$} if:
	\begin{enumerate}
		\item every arrow in $\mathcal{M}$ is a coproduct coprojection;
		\item $\X$ has all pushouts;
		\item $\X$ has strict initial objects and, for every object $X$, the arrow $?_X\colon 0\to X $ belongs to $\mathcal{M}$;
		\item all pushouts are $\mathcal{M}$-stable.
	\end{enumerate}
\end{definition}
